---
output: github_document
bibliography: vignettes/References.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/ReboundTools)](https://cran.r-project.org/package=EROITools)
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![Codecov test coverage](https://codecov.io/gh/earamendia/EROITools/branch/master/graph/badge.svg)](https://codecov.io/gh/earamendia/EROITools?branch=master)
<!-- [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.5841962.svg)](https://doi.org/10.5281/zenodo.5841962) -->
[![R-CMD-check](https://github.com/earamendia/EROITools/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/earamendia/EROITools/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->



# EROITools

The `R` package `EROITools` provides tools to aggregate Energy Return On Investment (EROI) values previously calculated using a Physical Supply Use Table (PSUT) framework to represent the Energy Conversion Chain (see @Heun:2018). Previous to using the `EROITools` package, the World Energy Extended Balances (WEEB) from the International Energy Agency (IEA) can be loaded and tidied using the `IEATools` and `ECCTools` `R` packages. The `Recca` `R` package then allows analysts to calculate a wide range of EROIs at the product and industry levels, and the `EROITools` package provides tools aggregate the calculated EROIs by product group. Important features of the `EROITools` package include the following:

* Aggregations can be performed either at the global or national level;
* Aggregations can be performed at both the primary and final stage of energy use, but also at the **useful** stage of energy use providing that final-to-useful efficiencies are provided by the analyst for each energy product;
* At the useful stage, the package also allows analysts to conduct these aggregations respecting sectoral or end-use breakdowns, so that the average useful stage EROI of a product group can be calculated for a given final demand sector (for instance steelmaking) or for a given end-use category (for instance high temperature heating);
* Last, the package allows analysts to add additional energy requirements (supposing that these are provided by the analyst) that cannot be quantified using the IEA's WEEB such as supply chain energy requirements --- see @brockwayEstimationGlobalFinalstage2019 or @brandcorreaDevelopingInputOutputBased2017 for examples of a quantification of such indirect energy flows.


## Installation

You can install `EROITools` from github with:

```{r gh-installation, eval = FALSE}
# install devtools if not already installed
# install.packages("devtools")
devtools::install_github("earamendia/EROITools")
# To build vignettes locally, use
devtools::install_github("earamendia/EROITools", build_vignettes = TRUE)
```


## History

This package builds upon the previous `IEATools`, `Recca`, and `ECCTools` `R` packages and will be demonstrated in a paper as soon as possible. The calculations conducted within the package are heavily dependent on the Physical Supply Use Table framework introduced to represent the Energy Conversion Chain in @Heun:2018 and further developed in @mr_psut_paper.


## More Information

Find more information, including vignettes and function documentation, at 
<https://earamendia.github.io/EROITools/>.

## References

