---
title: "Aggregating EROIs by fossil fuel group"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Aggregating EROIs by fossil fuel group}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: References.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(dplyr)
library(EROITools)
```


# Domestic perspective

## Primary and final stage aggregations

Primary stage EROIs can be simply aggregated with the `aggregate_primary_stage_erois`. The function calculates, for each fossil fuel group (all fossil fuels, coal products, natural gas, oil and gas products, and oil products), the shares of each product in the primary extraction phase (which iw why the `tidy_AB_dta` is passed as argument). Then, EROIs are computed by calculating the weighted average of EROIs. So, for country A, the share of coal in the group "Coal products" will be one, and the aggregated EROI for "Coal products" will be equal to the one of coal.

```{r}
# Aggregating primary stage EROIs:
aggregated_primary_stage_erois <- aggregate_primary_stage_erois(
    .tidy_erois_df = tidy_AB_erois_dta,
    .tidy_iea_df = tidy_AB_dta,
    eroi_calc_method = "dta"
  ) %>% 
  dplyr::glimpse()
```

Likewise, aggregated final stage EROIs are calculated with the `aggregate_final_stage_erois`. This function calculates, for each fossil fuel group, the shares of each product in the final use of a given country. It does so for the different final energy carriers: electricity, heat, fuel, and all carriers together. Then, aggregated final stage EROIs are calculated according to those shares.

```{r}
# Aggregating final stage EROIs:
aggregated_final_stage_erois <- aggregate_final_stage_erois(
    .tidy_erois_df = tidy_AB_erois_dta,
    .tidy_iea_df = tidy_AB_dta,
    eroi_calc_method = "dta"
  ) %>% 
  dplyr::glimpse()
```


## Useful stage aggregations

Aggregating EROIs at the useful stage of energy use requires the analyst to provide the average final-to-useful energy efficiencies with which each energy product is used, in each country. Indeed, the `Recca` package that we have used to calculate EROIs is used on a data frame describing the Energy Conversion Chain until the final energy stage, so that EROIs can only be calculated for products corresponding to the primary extraction and final energy stages.

Alternatively, the `Recca` package could also be used to calculate useful stage EROIs, but the matrices used should then describe the Energy Conversion Chain until the useful energy stage, which would require some preliminary work from the analyst anyway. To see example of how the Physical Supply Use Table and Input-Output matrices can be used to describe an Energy Conversion Chain until the useful energy stage, see @Heun:2018.

Here, we will create an artificial ad-hoc data frame containing made-up final-to-useful efficiencies to demonstrate the package.

```{r}
# Aggregating final stage EROIs:
# Pushing to tidy useful stage EROIs
length_to_use <- tidy_AB_erois_dta %>%
  dplyr::select(Country, Method, Energy.type, Year, Product) %>%
  dplyr::distinct() %>%
  nrow()

tidy_FU_efficiencies_dta <- tidy_AB_erois_dta %>%
  dplyr::select(Country, Method, Energy.type, Year, Product) %>%
  dplyr::distinct() %>%
  dplyr::mutate(
    Average_Efficiency_Col = seq(0.15, 1, 0.85/(length_to_use-1))
  ) %>% 
  dplyr::glimpse()
```

The analyst should provide the final-to-useful efficiencies in a format similar to the `tidy_FU_efficiencies_dta` data frame introduced here. Then, we can calculate the useful stage EROIs at the product level using the `push_to_useful_erois` as:

```{r}
# Binding efficiencies to Input-Output EROIs
tidy_useful_erois_dta <- push_to_useful_erois(
    .tidy_io_erois = tidy_AB_erois_dta,
    tidy_FU_efficiencies = tidy_FU_efficiencies_dta,
    eroi_calc_method = "dta"
  ) %>%
  dplyr::glimpse()
  #dplyr::filter(! is.na(Useful_Stage_EROI))
```

Finally, we can aggregate useful stage EROIs in a similar way than at the primary and final stages using the `aggregate_useful_stage_erois` function. The function also uses the final use shares of each product within each fossil fuel group, as for the `aggregate_final_stage_erois` function.

```{r}
# Aggregating final stage EROIs:
useful_stage_erois <- aggregate_useful_stage_erois(
  .tidy_erois_df = tidy_useful_erois_dta,
  .tidy_iea_df = tidy_AB_dta,
  eroi_calc_method = "dta"
) %>% 
  dplyr::glimpse()
```


# Multi-regional perspective

## Primary and final stage aggregations

To perform the aggregations when using a multi-regional perspective, the tidy data frame describing the Energy Conversion Chain needs to be adapted first, so that the country of location of each flow is added to the country column, and removed from the flow column. This step is important to allow the function to calculate the share of supply (primary stage) and use (final and useful stages) for each country. The `prepare_gma_for_shares` function is used to this effect.

```{r}
# Aggregating final stage EROIs:
tidy_AB_data_gma_prepared <- tidy_AB_data_gma %>%
  prepare_gma_for_shares() %>% 
  dplyr::glimpse()
```

Note the difference with the previous data frame:


```{r}
# Aggregating final stage EROIs:
dplyr::glimpse(tidy_AB_data_gma)
```

Then the aggregation is simply conducted as:

```{r}
# Aggregating final stage EROIs:
aggregated_primary_stage_erois_gma <- aggregate_primary_stage_erois(
    .tidy_erois_df = tidy_AB_erois_gma,
    .tidy_iea_df = tidy_AB_data_gma_prepared,
    eroi_calc_method = "gma"
  ) %>% 
  dplyr::glimpse()
```

and likewise, at the final stage, as:

```{r}
# Aggregating final stage EROIs:
aggregated_final_stage_erois_gma <- aggregate_final_stage_erois(
    .tidy_erois_df = tidy_AB_erois_gma,
    .tidy_iea_df = tidy_AB_data_gma_prepared,
    eroi_calc_method = "gma"
  ) %>% 
  dplyr::glimpse()
```


## Useful stage aggregations

As for the demonstration at the domestic level, we build a made-up data frame containing the average final-to-useful efficiencies of each energy product to expand the EROI calculations at the useful stage for each energy product.

```{r}
# Building a data frame containing average final-to-useful efficiencies for each energy product
length_to_use <- tidy_AB_erois_gma %>%
  dplyr::select(Country, Method, Energy.type, Year, Product) %>%
  dplyr::distinct() %>%
  nrow()

tidy_FU_efficiencies_gma <- tidy_AB_erois_gma %>% 
  dplyr::mutate(
    Country = stringr::str_extract(Product, "\\{.*\\}") %>% 
      stringr::str_remove("\\{") %>% stringr::str_remove("\\}"),
    Product = stringr::str_remove(Product, "\\{.*\\}_")
  ) %>% 
  dplyr::select(Country, Method, Energy.type, Year, Product) %>% 
  dplyr::distinct() %>% 
  dplyr::mutate(
    Average_Efficiency_Col = seq(0.15, 1, 0.85/(length_to_use-1))
  ) %>% 
  dplyr::glimpse()
```


Then, we expand the EROI calculations to the useful stage by applying the final-to-useful efficiencies of each energy product using the `push_to_useful_erois` function:

```{r}
# Pushing to tidy useful stage EROIs
tidy_useful_erois_gma <- push_to_useful_erois(
  .tidy_io_erois = tidy_AB_erois_gma,
  tidy_FU_efficiencies = tidy_FU_efficiencies_gma,
  eroi_calc_method = "gma"
) %>% 
  dplyr::glimpse()
```

Last, we can aggregate the useful stage EROIs using the `aggregate_useful_stage_erois` function:

```{r}
# Calculating aggregated useful stage EROIs:
aggregated_useful_stage_erois <- aggregate_useful_stage_erois(
  .tidy_erois_df = tidy_useful_erois_gma,
  .tidy_iea_df = tidy_AB_data_gma_prepared,
  eroi_calc_method = "gma"
) %>% 
  dplyr::glimpse()
```
