---
title: "Breaking down useful stage EROIs by sector or end-use"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Breaking down useful stage EROIs by sector or end-use}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: References.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(dplyr)
library(EROITools)
```


In this vignette, we show how EROIs can be aggregated by fossil fuel group at the useful stage by respecting a given breakdown, which would usually be a breakdown by final demand sector or end-use category.


# Domestic perspective

We start by constructing the dummy final-to-useful efficiencies for each energy product in a similar way than in the vignette demonstrating the aggregation process:

```{r}
# Constructing final-to-useful efficiencies data frame
length_to_use <- tidy_AB_erois_dta %>%
  dplyr::select(Country, Method, Energy.type, Year, Product) %>%
  dplyr::distinct() %>%
  nrow()

tidy_FU_efficiencies_dta <- tidy_AB_erois_dta %>%
  dplyr::select(Country, Method, Energy.type, Year, Product) %>%
  dplyr::distinct() %>%
  dplyr::mutate(
    Average_Efficiency_Col = seq(0.15, 1, 0.85/(length_to_use-1))
  ) %>% 
  dplyr::glimpse()
```

Now, supposing that the analyst has access to different final-to-useful efficiencies for three different end-uses; "heating", "machanical work", and "other end-uses"; let's suppose that the final-to-useful efficiencies for mechanical work are half of those of heating, and those of other end-uses a third of those of heating. We can then construct the data frame of end-use specific final-to-useful efficiencies as follows:

```{r}
# Making the data frame end-use specific
tidy_FU_efficiencies_dta_end_uses <- tidy_FU_efficiencies_dta %>% 
  tidyr::expand_grid(End_Use = c("Heating", "Mechanical Work", "Other end-uses")) %>% 
  dplyr::mutate(
    Average_Efficiency_Col = dplyr::case_when(
      Average_Efficiency_Col == "Mechanical Work" ~ Average_Efficiency_Col / 2,
      Average_Efficiency_Col == "Other end-uses" ~ Average_Efficiency_Col / 3,
      TRUE ~ Average_Efficiency_Col
    )
  ) %>% 
  dplyr::glimpse()
```

Then, we can calculate the useful stage EROIs at the product level using the `push_to_useful_erois`, which will respect the different end-uses.

```{r}
# Constructing useful stage EROIs data frame
  tidy_useful_erois_dta_end_uses <- push_to_useful_erois(
    .tidy_io_erois = tidy_AB_erois_dta,
    tidy_FU_efficiencies = tidy_FU_efficiencies_dta_end_uses,
    eroi_calc_method = "dta"
  ) %>% 
  dplyr::glimpse()
```

Note that likewise, a data frame containing the final demand sector specific final-to-useful efficiencies could have been conducted, and a final demand sector specific of useful stage EROIs could have been conducted.

Last, note that the `aggregate_useful_stage_erois` function does not yet allow to aggregate EROIs respecting the breakdown. Such a feature will be developed in the near future.


<!--
Finally, we can aggregate useful stage EROIs in a similar way than when no breakdown is provided, using the `aggregate_useful_stage_erois` function. The function also uses the final use shares of each product within each fossil fuel group, as for the `aggregate_final_stage_erois` function.
!-->



# Multi-regional perspective

Similarly, we construct the data frame containing the final-to-useful efficiencies.

```{r}
# Constructing final-to-useful efficiencies data frame
# length_to_use <- tidy_AB_erois_dta %>%
#   dplyr::select(Country, Method, Energy.type, Year, Product) %>%
#   dplyr::distinct() %>%
#   nrow()
# 
# tidy_FU_efficiencies_gma <- tidy_AB_erois_gma %>% glimpse()
#   dplyr::mutate(
#     Country = stringr::str_extract(Product, "\\{.*\\}") %>%
#       stringr::str_remove("\\{") %>% stringr::str_remove("\\}"),
#     Product = stringr::str_remove(Product, "\\{.*\\}_")
#   ) %>%
#   dplyr::select(Country, Method, Energy.type, Year, Product) %>%
#   dplyr::distinct() %>%
#   dplyr::mutate(
#     Average_Efficiency_Col = seq(0.15, 1, 0.85/(length_to_use-1))
#   ) %>%
#   dplyr::glimpse()
```


Next, we make this data frame end-use specific assuming some specific final-to-useful efficiencies by end-use: the average efficiency of mechanical work being half of the one of heating, and the one of other end-uses being a third of the one of heating, and this for all energy products.

```{r}
# Making the data frame end-use specific
# tidy_FU_efficiencies_gma_end_uses <- tidy_FU_efficiencies_gma %>% 
#   tidyr::expand_grid(End_Use = c("Heating", "Mechanical Work", "Other end-uses")) %>% 
#   dplyr::mutate(
#     Average_Efficiency_Col = dplyr::case_when(
#       Average_Efficiency_Col == "Mechanical Work" ~ Average_Efficiency_Col / 2,
#       Average_Efficiency_Col == "Other end-uses" ~ Average_Efficiency_Col / 3,
#       TRUE ~ Average_Efficiency_Col
#     )
#   ) %>% 
#   dplyr::glimpse()
```

Finally, we construct the useful stage EROIs data frame using the `push_to_useful_erois` function.

```{r}
# Constructing useful stage EROIs data frame
# tidy_useful_erois_gma_end_uses <- push_to_useful_erois(
#   .tidy_io_erois = tidy_AB_erois_gma,
#   tidy_FU_efficiencies = tidy_FU_efficiencies_gma_end_uses,
#   eroi_calc_method = "gma"
#   ) %>% 
#   dplyr::glimpse()
```
